
Library
```{r}
library(ggplot2)
library(dplyr)
```


```{r}
df <- read.csv('C:\\Users\\Kiersten\\Documents\\School\\Session_4\\Alistair\\MarketDataMonopoly.csv')
View(df)
summary(df)
```


```{r}
df$Price <- df$Revenue/df$Production #demand is endogenous
View(df)
```

***Define Q***
- Q as a function of two exogenous variables (input cost index and demand index)
```{r}
ggplot(df, aes(x=InputCostIndex, y=Production)) +
      geom_point()

```

```{r}
ggplot(df, aes(x=DemandIndex, y=Production)) +
      geom_point()
```

Clear linear relationship between input cost index and demand index
```{r}
r_quantity = lm(Production ~ InputCostIndex + DemandIndex, data=df)
summary(r_quantity)
# ask if input and demand indices changed, what would the optimal output be?
```

Plot production (q) as a function of input cost and demand indices
```{r}
f1 <- function(x,y) {
    2.523 -.09095*(x)+4.477*(y)
}

```

Build cost function
```{r}
df$Quantity <- f1(df$InputCostIndex, df$DemandIndex)
View(df)

r_cost = lm(Cost ~ InputCostIndex + Quantity, data=df)
summary(r_cost)

```

Plot production (q) as a function of input cost and demand indices
```{r}
f_cost <- function(x,y) {
    -307.9 + 1.544*(x)+19.84*(y)
}
```


```{r}
r_price = lm(Price ~ DemandIndex + Quantity, data=df)
summary(r_price)
```

```{r}
f_price <- function(x,y) {
    10.500652 + 2.944078*(x)-0.261707*(y)
}
```


```{r}
predict_function <- function(CostIndex, DemandIndex){
  quantity <- 2.523e+00 - 9.095e-02*CostIndex + 4.477e+00*DemandIndex
  cost <- -307.9 + 1.544*quantity + 19.84*CostIndex
  price <- 10.500652 + 2.944078*quantity -0.261707*DemandIndex
  profit <- quantity*price - cost
  
  output <- data.frame('Production' = quantity, 'Cost' = cost, 'Profit' = profit)
  print(output)
}
```

```{r}
predict_function(100,10)
predict_function(200,5)
predict_function(50,17)
```


---
NOTES FROM ALISTAIR:

partial derivatives (how does optimal q change)

Think through functional form

Given numerical optimization of the two functions we've chosen, what's the profit and cost

q is endogenous as a response to demand and input cost index (what is cost and demand function as they produce different q), control for endogeneity

- Solve for exogenous
take away effect of cost index from q, then see how price function changes in q

shock = input cost index and demand index
inverse demand function (price) = demand index and shock
cost function = chosen quantity and input cost index

how can calibrate model with given data, then predict with scenarios we haven't seen

Monopolistic competition
- quadratic cost function ()
- linear inverse demand function (a-q)
- add curvature: square, log, add elasticities
- Index: proportional to market cost (how do movements in index tell us about overall markets)




