
```{r}
install.packages('gtools')
library(gtools)
library(tidyr)
library(dplyr)
library(ggplot2)
```

Read in data
```{r}
inTeams <- read.csv('C:\\Users\\Kiersten\\Downloads\\spi_global_rankings.csv')
head(inTeams)
```
Mean parameters (and log mean)
```{r}
mean.off<-mean(inTeams$off)
mean.def<-mean(inTeams$def)
lmean.off<-mean(log(inTeams$off))
lmean.def<-mean(log(inTeams$def)) 
```

Subset to just Premier League
```{r}
premLeague<-subset(inTeams,league=="Barclays Premier League")
premTeams<-premLeague$name
rownames(premLeague)<-premTeams
df.prem<-premLeague[,c("off","def")]
df.prem
```

Offense parameter is expected goals when playing against an average team (uses a Poisson model for number of goals produced)
- alpha is the offense parameter
- delta is the defense parameter
```{r}
lmean.def<- log(mean(df.prem$def))
lmean.off<- log(mean(df.prem$off))               
df.prem["alpha"]<-log(df.prem["off"])-lmean.def #team's offense minus the mean defense (the larger the better)
df.prem["delta"]<-lmean.off-log(df.prem["def"]) #mean offense minus the team's defense (the larger the better because the smaller the original defense score the better)
head(df.prem)
# now expected goals for/conceded will match the offense and defense parameters
```

Lists of the parameters
```{r}
alphaList<-df.prem$alpha
deltaList<-df.prem$delta
names(alphaList)<-rownames(df.prem)
names(deltaList)<-rownames(df.prem)
```

Function for random poisson draw of scoreline for particular match
```{r}
draw.score<-function(team1,team2){
    c(
        rpois(1,exp(alphaList[team1]-deltaList[team2])),
  rpois(1,exp(alphaList[team2]-deltaList[team1]))
    )
}
draw.score("Liverpool","Arsenal")
```

All possible matches in PL:
*538 includes dynamics for when a match is more important to one team than another (qualifies them for a tournament etc.)
```{r}
# All possible matches in a season
allMatches<-permutations(20, 2, v=rownames(df.prem),repeats.allowed=FALSE)
colnames(allMatches)<-c("home","away")
head(allMatches,9)
rows = nrow(allMatches)
print(rows)
```

```{r}
# Example scores through the entire season
ScoresMatrix <- matrix(nrow=nrow(allMatches),  ncol=4)

for (i in 1:2000) {
  for (ii in 1:nrow(allMatches)  ) {
     ScoresMatrix[ii,1:2]=allMatches[ii,] # drawing random teams
     ScoresMatrix[ii,3:4]= draw.score(allMatches[ii,"home"],allMatches[ii,"away"] ) # simulating random score     lines (using defense/offense scores)
  }
}
  colnames(ScoresMatrix)<-c("home.team","away.team","home.score","away.score")
  head(ScoresMatrix)
```
What we have:
- All possible matches
- Simulation pulling all possible matches and simulating score lines of those matches

What we need:
- Each team playing each other twice (20 teams, each team plays 38 matches), 380 matches for one season
- League placement calculated based on season outcome: wins (3), draws (1), losses (0)
- Position earnings: convert league placement into position earnings (convert each position into monetary value)


```{r}
df.points <- as.data.frame(ScoresMatrix)

name <- rownames(df.prem)

for (i in 1:20){
    df.points[, name[i]] <- 0
}
```


```{r}
for (i in 1:380){
    for (j in 1:20){
        if (df.points[i,1] == name[j]){
            if (df.points[i,3] > df.points[i,4]){
                df.points[i,j+4] = 3
            } else if (df.points[i,3] == df.points[i,4]){
                df.points[i,j+4] = 1
            } else if (df.points[i,3] < df.points[i,4]){
                df.points[i,j+4] = 0
            } 
        } 
        if (df.points[i,2] == name[j]){
            if (df.points[i,3] > df.points[i,4]){
                df.points[i,j+4] = 0
            } else if (df.points[i,3] == df.points[i,4]){
                df.points[i,j+4] = 1
            } else if (df.points[i,3] < df.points[i,4]){
                df.points[i,j+4] = 3
            }      
        }
    }   
}
```


```{r}
ScoresMatrix <- as.data.frame(ScoresMatrix)
PointsMatrix <- as.data.frame(PointsMatrix)
PointsMatrix$home.points <- NA
PointsMatrix$away.points <- NA
df.prem$season_pts <- NA
df.prem$goals_for <- NA
df.prem$goals_against <- NA


for (i in 1:nrow(PointsMatrix)) {
  if (PointsMatrix[i,3] > PointsMatrix[i,4]) {
    PointsMatrix[i,5] = 3 
    PointsMatrix[i,6] = 0
    } else if (PointsMatrix[i,3] < PointsMatrix[i,4]) {
    PointsMatrix[i,5] = 0 
    PointsMatrix[i,6] = 3
    } else if (PointsMatrix[i,3] == PointsMatrix[i,4]) {
    PointsMatrix[i,5] = 1 
    PointsMatrix[i,6] = 1
  }
}
```

```{r}
# Add column by team
df.points_original <- as.data.frame(ScoresMatrix)
df.points_goalsfor <- as.data.frame(ScoresMatrix)
df.points_goalsagainst <- as.data.frame(ScoresMatrix)

name <- rownames(df.prem)
name_df <- as.data.frame(name)
name_df$goals_for=paste0(name_df$name,"_goalsfor")
name_df$goals_against=paste0(name_df$name,"_goalsagainst")
name_goalsfor <- subset(name_df, select = c("goals_for"))
name_goalsagainst <- subset(name_df, select = c("goals_against"))
goals_for <- as.character(name_goalsfor[,1])
goals_against <- as.character(name_goalsagainst[,1])

for (i in 1:20){
    df.points_original[, name[i]] <- 0
}
head(df.points_original)
df.points <- df.points_original

for (i in 1:20){
    df.points_goalsfor[, name[i]] <- 0
}

for (i in 1:20){
    df.points_goalsagainst[, name[i]] <- 0
}

# Add column goals for and goals against by team
```


```{r}
# POINTS PER MATCH
for (i in 1:380){
    for (j in 1:20){
        if (df.points[i,1] == name[j]){
            if (df.points[i,3] > df.points[i,4]){
                df.points[i,j+4] = 3
            } else if (df.points[i,3] == df.points[i,4]){
                df.points[i,j+4] = 1
            } else if (df.points[i,3] < df.points[i,4]){
                df.points[i,j+4] = 0
            } 
        } 
        if (df.points[i,2] == name[j]){
            if (df.points[i,3] > df.points[i,4]){
                df.points[i,j+4] = 0
            } else if (df.points[i,3] == df.points[i,4]){
                df.points[i,j+4] = 1
            } else if (df.points[i,3] < df.points[i,4]){
                df.points[i,j+4] = 3
            }      
        }
    }
}

# GOALS FOR
for (i in 1:380){
    for (j in 1:20){
  if (df.points_goalsfor[i,1] == name[j]){ # looking at home team
            df.points_goalsfor[i,j+4] = df.points_goalsfor[i,3]  # home points go to goals for
        }
        if (df.points_goalsfor[i,2] == name[j]){ # looking at away team
            df.points_goalsfor[i,j+4] = df.points_goalsfor[i,4] # away points go to goals for
        }
    }
}

# GOALS AGAINST
for (i in 1:380){
    for (j in 1:20){
  if (df.points_goalsagainst[i,1] == name[j]){ # looking at home team
            df.points_goalsagainst[i,j+4] = df.points_goalsagainst[i,4]  # home points go to goals for
        }
        if (df.points_goalsagainst[i,2] == name[j]){ # looking at away team
            df.points_goalsagainst[i,j+4] = df.points_goalsagainst[i,3] # away points go to goals for
        }
    }
}
```


```{r}
points <- df.points[,c(-1,-2,-3,-4)]
points <- sapply(points, as.numeric)
points_sum <- as.data.frame(colSums(points))

goalsfor <- df.points_goalsfor[,c(-1,-2,-3,-4)]
goalsfor <- sapply(goalsfor, as.numeric)
goalsfor_sum <- as.data.frame(colSums(goalsfor))

goalsagainst <- df.points_goalsagainst[,c(-1,-2,-3,-4)]
goalsagainst <- sapply(goalsagainst, as.numeric)
goalsagainst_sum <- as.data.frame(colSums(goalsagainst))
```


```{r}
df.teams <- cbind(points_sum, goalsfor_sum, goalsagainst_sum)
colnames(df.teams)[1] <- 'points'
colnames(df.teams)[2] <- 'goals_for'
colnames(df.teams)[3] <- 'goals_against'

df.teams <- df.teams %>%
    mutate(net_goals = goals_for - goals_against) %>%
    mutate(tie_breaker = runif(20, 0, 1))
```



